Name: Alex Lundin
Assignment: HW1

# 1
# The Boston data set contains values about residents from the suburb of Boston. The whole frame contains 506 rows and 14 columns. Some interesting metrics are medv which is the median value of owner-occupied homes and tax, with is the full values property tax.
```{r}
if(!require('MASS')){
  library('MASS')
}
data(Boston)
str(Boston)
```

# 2
```{r}
head(Boston, n=10)
tail(Boston, n=2)
dataFrameBoston<-Boston
dataFrameBoston[5,]
head(1:5, n=10)
```

# 3
```{r}
# Store Boston into a dataframe
# Store entire first column of data frame into crimeColumnAsVector
dataFrameBoston<-Boston
crimeColumnAsVector <- dataFrameBoston[,1]

# Find Mean.
result.mean <- mean(crimeColumnAsVector)
print(result.mean)

# Find Median.
result.median <- median(crimeColumnAsVector)
print(result.median)

# Find Range.
result.range <- range(crimeColumnAsVector)
print(result.range)
```
# 4
# This histogram represents the frequency of crime rates in each of the 506 towns of Boston. The most common crime rates have the largest bars. The most common crime rate is between 0 and 10

```{r}
hist(crimeColumnAsVector)
print(crimeColumnAsVector)
```

# 5
# This correlation of -0.3 is a weak negative linear relationship.
```{r}
# Store entire 14th column of data frame into medvColumnAsVector
medvColumnAsVector <- dataFrameBoston[,14]
cor(crimeColumnAsVector, medvColumnAsVector)
```

# 6
```{r}
# Store entire 6th column of data frame into roomColumnAsVector
roomColumnAsVector <- dataFrameBoston[,6]
```

# 7
```{r}
# Store entire fourth column of data frame into chasColumnAsVector
chasColumnAsVector <- dataFrameBoston[,4]
```

# 8
```{r}
# Store entire 9th column of data frame into radColumnAsVector
radColumnAsVector <- dataFrameBoston[,9]
```

# 9
```{r}

```

# 10
```{r}
# Store entire 13th column of data frame into lstatColumnAsVector
lstatColumnAsVector <- dataFrameBoston[,13]

# create new data frame from 4 vectors
dataFrameSummary <- c(crimeColumnAsVector)
dataFrameSummary <- c(roomColumnAsVector)
dataFrameSummary <- c(lstatColumnAsVector)
dataFrameSummary <- c(medvColumnAsVector)
print(dataFrameSummary)

# find neighborhood with highest median value
frameIndex <- which.max(medvColumnAsVector)
print(frameIndex)

dataFrameSummary[162,]

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
