Name: Alex Lundin
Assignment: HW1

# 1
# The Boston data set contains values about residents from the suburb of Boston. The whole frame contains 506 rows and 14 columns. Some interesting metrics are medv which is the median value of owner-occupied homes and tax, with is the full values property tax.
```{r}
if(!require('MASS')){
  library('MASS')
}
data(Boston)
str(Boston)
```

# 2
```{r}
head(Boston, n=10)
tail(Boston, n=2)
dataFrameBoston<-Boston
dataFrameBoston[5,]
head(1:5, n=10)
```

# 3
```{r}
# Store Boston into a dataframe
# Store entire first column of data frame into crimeColumnAsVector
dataFrameBoston<-Boston
crimeColumnAsVector <- dataFrameBoston[,1]

# Find Mean.
result.mean <- mean(crimeColumnAsVector)
print(result.mean)

# Find Median.
result.median <- median(crimeColumnAsVector)
print(result.median)

# Find Range.
result.range <- range(crimeColumnAsVector)
print(result.range)
```
# 4
# This histogram represents the frequency of crime rates in each of the 506 towns of Boston. The most common crime rates have the largest bars. The most common crime rate is between 0 and 10

```{r}
hist(crimeColumnAsVector)
print(crimeColumnAsVector)
```

# 5
# This correlation of -0.3 is a weak negative linear relationship.
# This means crime is loosely correlated to home value, for these particular cities.
```{r}
# Store entire 14th column of data frame into medvColumnAsVector
medvColumnAsVector <- dataFrameBoston[,14]
cor(crimeColumnAsVector, medvColumnAsVector)

```

# 6
# The correlation of .6953 shows a medium strength positive correlation between number of rooms and median house values.
# This means that more rooms, means most cost, for the most part. There are outliers and the graphs shows that.
```{r}
# Store entire 6th column of data frame into roomColumnAsVector
roomColumnAsVector <- dataFrameBoston[,6]

# store custom vectors to new dataFrame
dataFrameCustom <- data.frame(x=medvColumnAsVector, y=roomColumnAsVector)

par()              # view current settings
opar <- par()      # make a copy of current settings
par(col.lab="red") # red x and y labels 
plot(dataFrameCustom, xlab="Median Value", ylab="Number of Rooms")

par(opar)          # restore original settings
cor(roomColumnAsVector, medvColumnAsVector)

```

# 7
# The proximity to the Charles River (chas) does affect median house value.
# Yes chas is a factor in home price.
```{r}
# Store entire fourth column of data frame into chasColumnAsVector
chasColumnAsVector <- dataFrameBoston[,4]

# store custom vectors to new dataFrame
dataFrameCustom2 <- data.frame(x=chasColumnAsVector, y=medvColumnAsVector)

par()              # view current settings
opar <- par()      # make a copy of current settings
par(col.lab="red") # red x and y labels 
plot(dataFrameCustom2, xlab="Chas", ylab="Median Value")

# make chas a factor and plot again
chasColumnAsFactor <- factor(chasColumnAsVector)
dataFrameCustom3 <- data.frame(x=chasColumnAsFactor, y=medvColumnAsVector)

plot(dataFrameCustom3, xlab="Chas Factor", ylab="Median Value")

par(opar)          # restore original settings

```

# 8
# Rad is a integer number representing proximty to radial highways.
# Summary gives the min, 1 Quarter, median, mean , 3 quarter and max
# unique displays the unique (no repeated) entries
# 
```{r}
# Store entire 9th column of data frame into radColumnAsVector
radColumnAsVector <- dataFrameBoston[,9]
summary(radColumnAsVector)
unique(radColumnAsVector)
sum(radColumnAsVector = 24)
```

# 9
```{r}
far <- factor

func1 <- function(radColumnAsVector){
  # loop to traverse vector
  radColumnAsVector + seq_along(radColumnAsVector)
  
  # if structure to determine what to store in factor array
  if( radColumnAsVector[seq_along] != 24)(
     far <- TRUE
  )else(
     far <- FALSE
    )
  

}

# store custom vectors to new dataFrame
#dataFrameCustom4 <- data.frame(x=far, y=medvColumnAsVector)

#par()              # view current settings
#opar <- par()      # make a copy of current settings
#par(col.lab="red") # red x and y labels 
#plot(dataFrameCustom4, xlab="Far", ylab="Median Value")

#par(opar)          # restore original settings    



```

# 10
# This city has a crime value within the normal distribution, and other values to match in terms of house cost.
```{r}
# Store entire 13th column of data frame into lstatColumnAsVector
lstatColumnAsVector <- dataFrameBoston[,13]

# create new data frame from 4 vectors
dataFrameSummary <- c(crimeColumnAsVector)
dataFrameSummary <- c(roomColumnAsVector)
dataFrameSummary <- c(lstatColumnAsVector)
dataFrameSummary <- c(medvColumnAsVector)

# find neighborhood with highest median value
frameIndex <- which.max(medvColumnAsVector)
print(frameIndex)

rows <- 1
rows <- 6
rows <- 13
rows <- 14

Boston[c(frameIndex), ]

```
