Name: Alex Lundin
Assignment: Project 1

# Linear Regression

This data set is from the following URL:
https://archive.ics.uci.edu/ml/datasets/Heart+Disease

This study pulls various pieces of data from subjects to learn about factors related to heart disease.

14 attributes of the original 76 attributes were used in the processed data sets.

Complete attribute documentation: 
3 age: age in years 
4 sex: sex (1 = male; 0 = female) 
9 cp: chest pain type 
-- Value 1: typical angina 
-- Value 2: atypical angina 
-- Value 3: non-anginal pain 
-- Value 4: asymptomatic 
10 trestbps: resting blood pressure (in mm Hg on admission to the hospital) 
12 chol: serum cholestoral in mg/dl 
16 fbs: (fasting blood sugar > 120 mg/dl) (1 = true; 0 = false) 
19 restecg: resting electrocardiographic results 
32 thalach: maximum heart rate achieved 
38 exang: exercise induced angina (1 = yes; 0 = no) 
40 oldpeak = ST depression induced by exercise relative to rest 
41 slope: the slope of the peak exercise ST segment 
-- Value 1: upsloping 
-- Value 2: flat 
-- Value 3: downsloping 
44 ca: number of major vessels (0-3) colored by flourosopy 
51 thal: 3 = normal; 6 = fixed defect; 7 = reversable defect 
58 num: diagnosis of heart disease (angiographic disease status) 
-- Value 0: < 50% diameter narrowing 
-- Value 1: > 50% diameter narrowing 
(in any major vessel: attributes 59 through 68 are vessels) 

# load the project data

```{r}


# different file locations for the project
dataPathHomeComputer <- "C:\\Users\\Alex\\Desktop\\Screen-Cleaner\\GitHub\\UTDSummer2018\\CS-4375.0U2-Machine-Learning\\Projects\\project1\\data\\processed.hungarian.data"

dataPathSchoolComputer <- "H:\\GitHub\\UTDSummer2018\\CS-4375.0U2-Machine-Learning\\Projects\\project1\\data\\processed.hungarian.data"

# create the dataframe for heart disease
df_hd <- read.table(dataPathSchool)

# sets the column names of the data frame with colnames function
colnames(df_hd) <- c("age", "sex", "cp", "trestbps", "chol", "fbs", "restecg", "thalach", "exang", "oldpeak", "slope", "ca", "thal", "num")

# separate out the train and test dataframes following a similar naming convention of original frame

# Set random seed to ensure reproducibility of the shuffle.
set.seed(1234)

# shuffle the df_hd and store into a new df_hd frame
df_hd_numberOfRows <- nrow(df_hd)
shuf_df_hd <- df_hd[sample(df_hd_numberOfRows), ]

# set train_data df_hd with the train_data indices
train_data_indices <- 1:round(0.75 * df_hd_numberOfRows)
train_data <- shuf_df_hd[train_data_indices, ]
# set test_data df_hd with the test_data indices
test_data_indices <- (round(0.75 * df_hd_numberOfRows) + 1):df_hd_numberOfRows
test_data <- shuf_df_hd[test_data_indices, ]

```


# investigate the data with names, str, head, summary and cor
```{r}
# look at the names of the data frame
nameArray <- names(df_hd)
printString <- "The names of the columns are:"

# print a useful message with a compact version of the columns using str
print(printString)
str(nameArray)

# show first 6 instances of the frame
head(df_hd)

```

The 4 main columns that correlate highly with a diagonois of heart disease are:
9, 38, 40 and 41
cp, exang, oldpeak and slope
chest pain, exercise induced angina, ST depression induced by exercise relative to rest and slope of the peak exercise ST segment.
all four of these have a correlation above .5 with the diagnois of heart disease in patients.
```{r}
# store a summary of the data frame
sm <- summary(df_hd)

# print the summary
print(sm)

# coerce all predictors and targets as numeric for correlation function

xPredictor <- as.numeric(df_hd$cp)
yTarget <- as.numeric(df_hd$num)

print("Correlation of -- Chest pain and heart disease:")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_hd$exang)
print("Correlation of -- Exercise induced chest pain and heart disease:")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_hd$oldpeak)
print("Correlation of -- ST depression, aka irregular heart beat, due to exercise and heart disease:")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_hd$slope)
print("Correlation of -- Slope of ST depression during peak exercise and heart disease:")
cor(xPredictor, yTarget)
```


# create first model for heart disease
# This model uses exercise induced angina as the predictor for the target of a heart disease diagnosis
The model has:
Rsquared = .38
low pvalues for the predictor and target
low pvalue for the f-stastic
```{r}
xPredictor <- as.numeric(train_data$exang)
yTarget <- as.numeric(train_data$num)
# create linear model on train_data
train_lm1 <- lm(xPredictor ~ yTarget, data=train_data)
print(train_lm1 )
plot(train_lm1)

lm1_sm <- summary(train_lm1)
print(lm1_sm)
```

```{r}
xPredictor_hp <- train_data$horsepower
yResponse_mpg <- train_data$mpg
# create linear model on train_data
train_linearModel <- lm(xPredictor_hp ~ yResponse_mpg, data=train_data)
print(train_linearModel)
plot(train_linearModel)
```