---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

Name: Alex Lundin
Assignment: HM4




1.	Set up the Auto data:
a.	Load the ISLR package and the Auto data
b.	Determine the median value for mpg
c.	Use the median to create a new column in the data set named mpglevel, which is 1 if mpg>median and otherwise is 0. 
Make sure this variable is a factor. We will use mpglevel as the target (response) variable for the algorithms. 
d.	Use the names() function to verify that your new column is in Auto
e.	create a 75-25 train/test split, using seed 1234 but do not include columns 'name' or 'mpg' in either train or test
```{r}
# a
if(!require('ISLR')){
  install.packages('ISLR')  
  library('ISLR')
}

# b

medianMPG <- median(Auto$mpg)
paste("The median MPG is: ", medianMPG)

# c
Auto$mpglevel <- as.factor(ifelse(Auto$mpg > medianMPG, 1, 0))
str(Auto$mpglevel)

# d
names(Auto)

# e
set.seed(1234)
i <- sample(1:nrow(Auto), 0.75*nrow(Auto))
train <- Auto[i,c(2,3,4,5,6,7,8,10)]
test <- Auto[-i,c(2,3,4,5,6,7,8,10)]

```




2.	Plots 
a.	 Set up a 2x2 graph grid and plot the following pairs of plots 
b.	Plot pair 1: plot horsepower~mpg and weight~mpg, setting colors according to the factor mpglevel, ex: col=(Auto$mpglevel)
c.	Plot pair 2:  plot horsepower~mpglevel and weight~mpglevel 
```{r}
# a
par(mfrow=c(2, 1))  # divide graph area in 2 columns
palette(c("red","black"))

# b
plot(Auto$horsepower~Auto$mpg, col=ifelse(Auto$mpglevel==1, "red", "black"))
plot(Auto$weight~Auto$mpg, col=ifelse(Auto$mpglevel==1, "red", "black"))

# c
plot(Auto$horsepower~Auto$mpglevel)
plot(Auto$weight~Auto$mpglevel)
```





3.	Build a Naïve Bayes model 
a.	build the model on the train set
b.	use the predict() function on the test set
c.	create a table comparing predicted to actual values for mpglevel
d.	calculate the mean accuracy
      0.928571428571429

```{r}

# a
library(e1071)
nb1 <- naiveBayes(mpglevel~., data=train)

# b
p1 <- predict(nb1, newdata=test, type="class")

# c
table(p1, test$mpglevel)

# d
paste("The mean accuracy is:", mean(p1==test$mpglevel))


```





4.	SVM linear kernel 
a.	use the tune() function to perform cross-validation to determine the best value for cost
b.	use the parameter(s) from the previous step to build an svm model with a linear kernel on the train set
c.	use the predict() function on the test set
d.	create a table comparing predicted to actual values for mpglevel
e.	calculate the mean accuracy
      The mean accuracy is: 0.908163265306122
      
```{r}

library(e1071)
svm_fit1 <- svm(mpglevel~., data=train, kernel="linear", cost=10, scale=FALSE)

# a
tune_svm1 <- tune(svm, mpglevel~., data=train, kernel="linear",
               ranges=list(cost=c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
summary(tune_svm1)

# b
best_mod1 <- tune_svm1$best.model
summary(best_mod1)

# c
svm_pred <- predict(best_mod1, newdata=test)

# d
table(svm_pred, test$mpglevel)

# e
paste("The mean accuracy is:", mean(svm_pred==test$mpglevel))

```





5.	SVM polynomial kernel 
a.	use the tune() function to perform cross-validation to determine the best values for cost and degree
b.	use the parameter(s) from the previous step to build an svm model with a polynomial kernel on the train set
c.	use the predict() function on the test set
d.	create a table comparing predicted to actual values for mpglevel
e.	calculate the mean accuracy





6.	SVM radial kernel 
a.	use the tune() function to perform cross-validation to determine the best values for cost and gamma
b.	use the parameter(s) from the previous step to build an svm model with a radial kernel on the train set
c.	use the predict() function on the test set
d.	create a table comparing predicted to actual values for mpglevel
e.	calculate the mean accuracy
      The mean accuracy is: 0.928571428571429
```{r}

library(e1071)
svm_fit3 <- svm(mpglevel~., data=train, kernel="radial", cost=1, gamma=1, scale=FALSE)

# a
tune_svm3 <- tune(svm, mpglevel~., data=train, kernel="radial",
               ranges=list(cost=c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
summary(tune_svm3)

# b
best_mod3 <- tune_svm3$best.model
summary(best_mod3)

# c
svm_pred3 <- predict(best_mod3, newdata=test)

# d
table(svm_pred3, test$mpglevel)

# e
paste("The mean accuracy is:", mean(svm_pred3==test$mpglevel))


```





7.	Questions. 
a.	Compare the accuracy results for the 4 models
b.	Discuss the advantages and disadvantages of Naïve Bayes versus svm
