---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

Name: Alex Lundin
Assignment: Project 2
# Header for project

Data URL:
https://www.kaggle.com/new-york-state/nys-traffic-tickets-issued-four-year-window

Data Overview:

Data Label 			Data Type 					Data Description


VIOL_CHRGD 			Text						Law code of the violation for which the ticket was issued.
        												Mostly comports with NYS Vehicle & Traffic Law, but can include codes
        												from NYC Traffic Regulations, Thruway Regulations, Parkway
        												Regulations, Penal law, Transportation Law and Tax Law.
        												May also include administrative codes attached to the end of the law
        												code that allow courts to differentiate between various actions covered
        												by a single section of law (for instance 1180D1A, where the "A" refers
        												to a speed in zone violation between 11-30 MPH over the posted speed
        												limit).
												
VIOL_DESC 			Text 						Description of the violation for which the ticket was issued.

VIOL_YR 			  Integer 			  Calendar year in which the ticket was issued.

VIOL_MO 			  Integer 				Month in which the ticket was issued.

VIOL_DAY 			  Text 						Day of the week on which the ticket was issued.

AGE_AT_VIOL 		Integer 			  Age, in years, of the person to whom the ticket was issued on the date
												        of ticket issuance.

STATE_OF_LIC 		Text						Name of the state, district, territory, province or country that issued the
												        driver license presented by the motorist at the time of ticket issuance.

GENDER 				  Text						Gender of person to whom ticket was issued.
												        M = Male ; F = Female; C = Organization; U = Unknown

POLICE_AGENCY 	Text 					  The name of the law enforcement agency that issued the ticket.

COURT 				  Text 						Court where the ticket was answerable.

SOURCE 				  Text						Processing system used by the courts to record and track tickets from
												        issuance to disposition:

												        TSLED = "Traffic Safety Law Enforcement & Disposition Program" used
												        by local courts throughout most of NYS.

												        TVB = used by NYSDMV Traffic Violation Bureaus in NYC, western
												        Suffolk County, Rochester and Buffalo (Suffolk County and Buffalo TVB
												        offices closed on 4/1/2013 and 7/1/2015, respectively).
												        
```{r}
# store path string for loading csv file
dataPathHomeComputer <- "C:\\Users\\Alex\\Desktop\\data\\traffic-tickets-issued-four-year-window.csv"

df_tt <- read.csv(dataPathHomeComputer)
```
	
```{r}
# separate out the train and test dataframes following a similar naming convention of original frame

# Set random seed to ensure reproducibility of the shuffle.
set.seed(1958)


 
# Split data in half, exactly
train_data = df_tt[1:(0.001*nrow(df_tt)),]
dim(train_data)  # 6 11
test_data = df_tt[((0.001*nrow(df_tt))+1):((0.001*nrow(df_tt)*2)),]
dim(test_data) # 26 11

```



# investigate the data with names, str, head, summary and cor
```{r}
# show how many rows
paste("There are over 500,000 observances, also known as rows.")
str(df_tt)

# look at the names of the data frame
nameArray <- names(df_tt)

# print a useful message with a compact version of the columns
paste("The names of the columns are:")
paste("Column:", nameArray)

# show first 6 instances of the frame
head(df_tt)

```

The 4 main columns that correlate highly with type of violation are:

```{r}
# store a summary of the data frame
sm <- summary(df_tt)

# print the summary
print(sm)

# coerce all predictors and targets as numeric for correlation function

xPredictor <- as.numeric(df_tt$Age.at.Violation)
yTarget <- as.numeric(df_tt$Violation.Charged.Code)

print("Correlation of -- Age and violation code:")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_tt$Police.Agency)
print("Correlation of -- Police Agency and violation code:")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_tt$State.of.License)
print("Correlation of -- State of License and violation code::")
cor(xPredictor, yTarget)

xPredictor <- as.numeric(df_tt$Violation.Year)
print("Correlation of -- Violation Year and violation code:")
cor(xPredictor, yTarget)

corMatrix <- (cor(df_tt[,2:3]))
findCorrelation(corMatrix, cutoff=0.5, verbose=TRUE)

```

# two informative graphs
<!-- The scatter plot shows the likelyhood of heart disease for each type of chest pain (which ranges from 1-4) -->
<!-- The scatter plot function in R complains when there is a fixed number of observance values (chest pain from 1-4) -->

<!-- The density plots show the highest correlators do not have a normal distrubtion of occcurances in the data set. -->
<!-- The predictors are very un normally distributed. -->
<!-- ```{r} -->
<!-- # scatterplot for data view -->
<!-- scatter.smooth(x=df_tt$cp, y=df_tt$num, main="Scatterplot Chest Pain ~ Heart Disease", xlab="Chest Pain (types 1 through 4)", ylab="Heart disease (< .5 positive diagnosis)") -->


<!-- # # Density plot -->
<!-- par(mfrow=c(1, 2))  # divide graph area in 2 columns -->

<!-- plot(density(df_tt$cp), main="Density Plot: Chest Pain", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(df_tt$cp), 2))) -->
<!-- polygon(density(df_tt$cp), col="red") -->

<!-- plot(density(df_tt$exang), main="Density Plot: Exercise induced CP", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(df_tt$exang), 2))) -->
<!-- polygon(density(df_tt$exang), col="red") -->


<!-- ``` -->


# first model for traffic ticket data
# This model uses state of drivers liscense as the predictor for the target of violation charged code
The model has:
Rsquared = .012
low pvalues for the predictor and target
low pvalue for the f-stastic

The rsquared value shows that this type of model performs bad on the data

I think the prediction is bad because individually, none of the predictors correlate with the target
```{r}
# create linear model on train_data
train_lm1 <- lm(as.numeric(train_data$State.of.License)~as.numeric(train_data$Violation.Charged.Code), data=train_data)
# print(train_lm1 )
# plot(train_lm1)

# store a summary of the model created from the train data and print it
lm1_sm <- summary(train_lm1)
print(lm1_sm)

# attempt to predict the target y value with the train linear model
pred <- predict(train_lm1, test_data)
print("Correlation of -- Prediction of violation code:")
cor(pred, as.numeric(test_data$State.of.License),use='complete.obs')
```

# second model for traffic ticket data
# naiveBayes

```{r}

# a
library(e1071)
nb1 <- naiveBayes(as.numeric(train_data$Age.at.Violation)~as.numeric(train_data$Violation.Charged.Code), data=train_data)
# 
# # b
p1 <- predict(nb1, newdata=test_data)

# c
table(p1, test_data$Violation.Charged.Code, type="class")

# d
paste("The mean accuracy is:", mean(p1==test_data$Violation.Charged.Code))


```

# third model for traffic ticket data
# SVM radial kernel 
```{r}

svm_fit3 <- svm(train_data$Violation.Charged.Code~train_data$Age.at.Violation, data=train_data, kernel="radial", cost=1, gamma=1, scale=FALSE)

# a
tune_svm3 <- tune(svm, train_data$Violation.Charged.Code~., data=train_data, kernel="radial",
                ranges=list(cost=c(0.1,1,10,100,1000),
                gamma=c(0.5,1,2,3,4)))
summary(tune_svm3)

# b
best_mod3 <- tune_svm3$best.model
summary(best_mod3)

# c
svm_pred3 <- predict(best_mod3, newdata=test_data)

# d
table(svm_pred3, test_data$Violation.Charged.Codel)

# e
paste("The mean accuracy is:", mean(svm_pred3==test_data$Violation.Charged.Code))


```

